\documentclass{article}

\usepackage{extramarks}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

\usepackage{hyperref}

\usepackage{tikz}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[shortlabels]{enumitem}

\usepackage{float,graphicx}

\usepackage{pgfplots}

\usepackage{adjustbox}


\usepackage{array}
\usepackage{pgfgantt}

\usepackage[utf8]{inputenc}

\usepackage{fancyhdr}
\usepackage{fancybox}

\topmargin=-0.45in
\evensidemargin=0in
\oddsidemargin=0in
\textwidth=6.5in
\textheight=9.0in
\headsep=0.25in

% Listings' Styles

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.96,0.96,0.96}

\lstdefinestyle{python}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
 
\lstset{style=python}
\lstset{language=Python}

\linespread{1.1}
\captionsetup[table]{position=bottom}

\pagestyle{fancy}
\lhead{\groupName}
\chead{\hmwkClass: \hmwkTitle}
\rhead{\today}
\lfoot{\lastxmark}
\cfoot{\thepage}

\renewcommand\headrulewidth{0.4pt}
\renewcommand\footrulewidth{0.4pt}
\renewcommand\qedsymbol{$\blacksquare$}

\setlength\parindent{0pt}

\newcommand{\hmwkTitle}{Assignment Sheet\ \#\hmwkNumber}
\newcommand{\hmwkDueDate}{February 15, 2022}
\newcommand{\hmwkClass}{Machine Learning}
\newcommand{\hmwkAuthorName}{Henri Sota, Enis Mustafaj}
\newcommand{\groupName}{\textbf{Group HB}}

% Homework Number Variable
\newcommand{\hmwkNumber}{1}

%
% Create Problem Sections
%

\newcommand{\enterProblemHeader}[1]{
	\nobreak{}
}

\newcommand{\exitProblemHeader}[1]{
    \stepcounter{#1}
}

\setcounter{secnumdepth}{0}
\newcounter{partCounter}
\newcounter{homeworkProblemCounter}

\setcounter{homeworkProblemCounter}{1}
\nobreak\extramarks{Problem \hmwkNumber \arabic{homeworkProblemCounter}}{}\nobreak{}

%
% Homework Problem Environment
%
% This environment takes an optional argument. When given, it will adjust the
% problem counter. This is useful for when the problems given for your
% assignment aren't sequential. See the last 3 problems of this template for an
% example.
%
\newenvironment{homeworkProblem}[1][-1]{
    \ifnum#1>0
        \setcounter{homeworkProblemCounter}{\hmwkNumber.#1}
    \fi
    \section{Problem \hmwkNumber.\arabic{homeworkProblemCounter}}
    \setcounter{partCounter}{1}
    \enterProblemHeader{homeworkProblemCounter}
}{
    \exitProblemHeader{homeworkProblemCounter}
}


\newcounter{programmingPartCounter}
\newcounter{programmingProblemCounter}

\setcounter{programmingProblemCounter}{1}
\nobreak\extramarks{Programming Problem \hmwkNumber \arabic{programmingProblemCounter}}{}\nobreak{}

%
% Programming Problem Environment
%
% This environment takes an optional argument. When given, it will adjust the
% problem counter. This is useful for when the problems given for your
% assignment aren't sequential. See the last 3 problems of this template for an
% example.
%
\newenvironment{programmingProblem}[1][-1]{
    \ifnum#1>0
        \setcounter{programmingProblemCounter}{\hmwkNumber.#1}
    \fi
    \section{Programming Problem \hmwkNumber.\arabic{programmingProblemCounter}}
    \setcounter{programmingPartCounter}{1}
    \enterProblemHeader{programmingProblemCounter}
}{
    \exitProblemHeader{programmingProblemCounter}
}

\title{
    \vspace{2in}
    \textmd{\textbf{\hmwkClass:\\ \hmwkTitle}}\\
    \normalsize\vspace{0.1in}\small{Due\ on\ \hmwkDueDate\ at 10:00}\\
    \vspace{3in}
}

\author{\groupName \\ \hmwkAuthorName}
\date{}

\renewcommand{\part}[1]{\textbf{\large Part \Alph{partCounter}}\stepcounter{partCounter}\\}

% Useful for algorithms
\newcommand{\alg}[1]{\textsc{\bfseries \footnotesize #1}}

% Alias for the Solution section header
\newcommand{\solution}{\textbf{\large Solution}}
\newcommand{\comment}[1]{} % Multi-line comment


\begin{document}
\maketitle
\pagebreak
\begin{homeworkProblem} 
The disjoint events $A$, $B$, and $C$ are defined in a sample space $\Omega$. Find expressions for the following probabilities in terms of sums of the probabilities $P(\emptyset)$, $P(A)$, $P(B)$, $P(A \cup B)$, $P(A \cup C)$, $P(B \cup C)$, $P(A \cup B \cup C)$:

\begin{enumerate}[a)]
    \item the probability that exactly  two of $A$, $B$, $C$ occur (\textit{at the same time!})
    \item the probability that exactly one of these events occurs
\end{enumerate}
\end{homeworkProblem}

\begin{homeworkProblem}
We consider the chance experiment of four tosses of a fair coin.

\begin{enumerate}[a)]
    \item Give the sample space $\Omega$ for this experiment and define a discrete random variable $X$ to describe the number of heads obained in four tosses of a fair coin.
    \item Find the PMF and CDF of the defined random variable and plot the CDF.
    \item Compute the probability of the event $X$ IS BIGGER THAN $1$.
    \item Compute the mean of the random variable.
\end{enumerate}
\end{homeworkProblem}

\begin{homeworkProblem}
Let the function $\rho$ with $\rho(x) = 
\begin{cases}
    c|x| (1 + x)(1 - x) & \text{if } -1 \leq x \leq 1\\
    0 & \text{else}
\end{cases}$
be given.

\begin{enumerate}[a)]
    \item Compute the constant $c$, such that $\rho$ becomes a density of some random variable.
    \item Consider a random variable $X$ with the just computed density $\rho$. Find its CDF and plot it.
    \item Use the CDF to compute \begin{equation*}
        P(X < -0.5), \quad P(X > 0.5), \quad P(-0.5 < X < 0.5)
    \end{equation*}
\end{enumerate}
\end{homeworkProblem}

\begin{programmingProblem}
In this first programming exercise, we would like to familiarize us with typical linear algebra computing tasks. To this end, we define the vectors / matrices

\begin{equation*}
    \mathbf{a} = \begin{pmatrix} 1 \\ 1 \\ 0 \end{pmatrix}, \quad
    \mathbf{b} = \begin{pmatrix} -1 \\ 2 \\ 5 \end{pmatrix}, \quad
    M = \begin{pmatrix} 2 & -1 & 0 \\ -1 & 2 & -1 \\ 0 & -1 & 2 \end{pmatrix}
\end{equation*}

\begin{enumerate}[a)]
    \item Compute and print the inner product between $\mathbf{a}$ and $\mathbf{b}$.
    \item Compute and print the matrix-vector product between $M$ and $\mathbf{b}$.
    \item Compute and print the $l_{2}$ norm of $\mathbf{b}$.
    \item Compute and print the solution $\mathbf{x}$ of the linear system of equations $M \mathbf{x} = \mathbf{b}$ using the LU factorization.
    \item Compute and print the solution $\mathbf{x}$ of the linear system of equations $M \mathbf{x} = \mathbf{b}$ using the Cholesky factorization.
\end{enumerate}
\end{programmingProblem}

\end{document}
